\chapter*{Acknowledgements}
\markboth{Acknowledgements}{}
\justifying
I would like to express my sincere gratitude to Indian Institute of Technology Gandhinagar for the past 5 years of my tenure as a PhD student in the institute. I would be failing in my duty, if I do not acknowledge the people who played a part in shaping this thesis.

First and Foremeost, I would like to thank my supervisor, \textbf{Prof. Sairam S. Mallajosyula} for initiating and teaching me the ways of academic research. You served as an example, teaching me the ways of managing the lab and research effectively. I would be forever indebted for the lessons I learned in the lab, and would like to express my heartfelt gratitude for the same.

I would like to thank my Doctoral Student Committee members, \textbf{Prof. Sriram Kanvah Gundimeda} and \textbf{Prof. Mithun Radhakrishna} for their valuble insights and questions, which offered a new perspective to my research directions. I would also like to thank \textbf{Prof. Gopinadhan Kalon} for the many discussions we had related to multiple topics in materials, and for teaching me the fundamentals of 2D nanomaterials. I am also grateful to \textbf{Prof. Anirban Mondal} for critical insights into my work.

I would like to thank my fiancee \textbf{Ms. Chythra J. N.} for her unconditional support and love during the testing periods. I would have been soon lost in this myriad of works, but you kept me tethered to the ground and to reality. It should come as no surprise that you played a large part in my overall growth as a person, and in the completion of this thesis. This thesis is the result of our combined blood, sweat and tears.

I am extremely grateful to \textbf{Dr. Richa Sharma}, Program Manager for CO\textsubscript{2} Capture Chemistry \& Measurements at Schlumberger-Doll Research (SDR) for providing me an oppurtunity to work as an intern with SDR. I am also grateful to \textbf{Prof. Nike Dattani} for building a warm and welcoming community at Matter Modelling Stack Exchange.

Finally, I would like to thank my fellow labmates and colleagues. Many thanks are due to Dr. Lata Rani for her help during my stay here. I would also like to acknowledge Dr. Neethish M.M, a colleague from Pondichery University for valuable support during my academic journey. I would like to thank Ms. Haritha "Oppol" Dileep, Mr. Pranav Umesh Bhagwat, Ms. Collinica "Khun" Camillie Syiemleih, Mr. Akhil A.R, Mrs. Deepshikha "Ghosh" Ghosh, Mr. Akshay Rajeev, Mr. Ajay "Flood" Mohan, Ms. Shikha Dhakkar and Mr. Pradeep "Tiger" Kumar Yadav for being the best friends that someone could ask for. Many thanks are due to friends from other disciplines Dr. Manu Kurian, Dr. Asha Liza James, Dr. Sethulakshmi N. and Prasanna "Kulku" Kulkarni for their support and making my time here enjoyable. I am also thankful to Dr. Simanta Lakhar and Dr. Kalyani Patrikar for entertaining my random questions and being a good sport.

Last, much gratitude is due to my parents, in-laws and siblings for their warmth and prayers. Above all, unlimited gratefulness is offered to the divine hand for the blessings showered on me, without which, nothing is possible.
\vspace{2em}
\begin{flushright}
    (Hemanth.H)
\end{flushright}